---
title: "Uber Trips Data Visualization"
author: "Drucila LeFevre"
output: html_document
---
#### Source : https://data-flair.training/blogs/r-data-science-project-uber-data-analysis/

#### Objective: Implement R visualization tools to gain insights about the Uber Pickups dataset


#### Observation: An increase in pickup traffics occurs in the evening from 5 to 6 pm. 
```{r warning=FALSE, message = FALSE, echo= FALSE}
library(knitr)
library(ggplot2)
library(ggthemes)
library(lubridate)
library(dplyr)
library(tidyr)
library(DT)
library(scales)
blues = '#003f5c'
colorsp = c("#003f5c", "#374c80", "#7a5195", "#bc5090", "#ef5675", "#ff764a", '#ffa600' )
uber = read.csv('uber.csv')
hour_data <- uber %>%
    group_by(hour) %>%
    dplyr::summarize(Total = n()) 

ggplot(hour_data, aes(hour, Total)) + 
    geom_bar( stat = "identity", fill = 'navy', color = "gray7") +
    ggtitle("Trips Every Hour") +
    theme(legend.position = "none") +
    scale_y_continuous(labels = comma)
```

```{r warning=FALSE, message = FALSE, echo= FALSE}
month_hour <- uber %>%
    group_by(month, hour) %>%
    dplyr::summarize(Total = n())

ggplot(month_hour, aes(hour, Total, fill = month)) +
    geom_bar( stat = "identity") +
    ggtitle("Trips by Hour and Month") +
    scale_y_continuous(labels = comma) + 
    scale_fill_manual(values = colorsp)
```

#### Observation: The highest number of trips occured on the 30th day of April 

```{r warning=FALSE, message = FALSE, echo= FALSE}
day_group <- uber %>%
    group_by(day) %>%
    dplyr::summarize(Total = n()) 



ggplot(day_group, aes(day, Total)) + 
    geom_bar( stat = "identity", fill = "navy") +
    ggtitle("Trips Every Day") +
    theme(legend.position = "none") +
    scale_y_continuous(labels = comma) + 
    scale_fill_manual(values = colorsp)
```

```{r warning=FALSE, message = FALSE, echo= FALSE}
day_month_group <- uber %>%
    group_by(month, day) %>%
    dplyr::summarize(Total = n())

ggplot(day_month_group, aes(day, Total, fill = month)) + 
    geom_bar( stat = "identity") +
    ggtitle("Trips by Day and Month") +
    scale_y_continuous(labels = comma) +
    scale_fill_manual(values = colorsp)

```

#### Observation: Highest number of trips were taken in the month of September. 


```{r warning=FALSE, message = FALSE, echo= FALSE}

month_group <- uber %>%
    group_by(month) %>%
    dplyr::summarize(Total = n()) 

ggplot( month_group, aes(month, Total, fill = month)) + 
    geom_bar( stat = "identity") +
    ggtitle("Trips by Month") +
    theme(legend.position = "none") +
    scale_y_continuous(labels = comma) +
    scale_fill_manual(values = colorsp)
```

#### Observation:In September, the highest number of trips were on Tuesday, Friday, and Saturday.
```{r warning=FALSE, message = FALSE, echo= FALSE}
month_weekday <- uber %>%
    group_by(month, dayofweek) %>%
    dplyr::summarize(Total = n())

ggplot(month_weekday, aes(month, Total, fill = dayofweek)) + 
    geom_bar( stat = "identity", position = "dodge") +
    ggtitle("Trips by Day and Month") +
    scale_y_continuous(labels = comma) +
    scale_fill_manual(values = colorsp)

```

#### Observation: In regards to the five bases in New York City, Base 02617 had the highest number of trips.


#### Base is the the TLC base company code affiliated with the Uber pickup. 
```{r warning=FALSE, message = FALSE, echo= FALSE}
ggplot(uber, aes(Base)) + 
    geom_bar(fill = "steelblue4") +
    scale_y_continuous(labels = comma) +
    ggtitle("Trips by Bases") 
```

#### Observation: Looking into the same base (B012617), its highest number of trips occured in September. 
```{r warning=FALSE, message = FALSE, echo= FALSE}
ggplot(uber, aes(Base, fill = month)) + 
    geom_bar(position = "dodge") +
    scale_y_continuous(labels = comma) +
    ggtitle("Trips by Bases and Month") +
    scale_fill_manual(values = colorsp)

```

#### Below are plots that delineates by day, hour, and months. 
```{r warning=FALSE, message = FALSE, echo= FALSE}
day_and_hour <- uber %>%
    group_by(day, hour) %>%
    dplyr::summarize(Total = n())

ggplot(day_and_hour, aes(day, hour, fill = Total)) +
    geom_tile(color = "white") +
    ggtitle("Heat Map by Hour and Day")

```

```{r warning=FALSE, message = FALSE, echo= FALSE}
ggplot(day_month_group, aes(day, month, fill = Total)) +
    geom_tile(color = "white") +
    ggtitle("Heat Map by Month and Day")
```

```{r warning=FALSE, message = FALSE, echo= FALSE}
ggplot(month_weekday, aes(dayofweek, month, fill = Total)) +
    geom_tile(color = "white") +
    ggtitle("Heat Map by Month and Day of Week")
```

```{r warning=FALSE, message = FALSE, echo= FALSE}
month_base <-  uber %>%
    group_by(Base, month) %>%
    dplyr::summarize(Total = n()) 

day0fweek_bases <- uber %>%
    group_by(Base, dayofweek) %>%
    dplyr::summarize(Total = n()) 

ggplot(month_base, aes(Base, month, fill = Total)) +
    geom_tile(color = "white") +
    ggtitle("Heat Map by Month and Bases")

```

```{r warning=FALSE, message = FALSE, echo= FALSE}
ggplot(day0fweek_bases, aes(Base, dayofweek, fill = Total)) +
    geom_tile(color = "white") +
    ggtitle("Heat Map by Bases and Day of Week")
```






